### 2014-11-13 tex/evaluation.tex
Finished Eval

%\begin{figure*}[t]
\begin{table*} [t]
  \centering
  \scalebox{0.88}{\begin{tabular}{| c | c | r | r | r | r | r | r | r | r | r |}
  \hline
   \multirow{2}{*}{Method }&\multirow{2}{*}{ $k$ }
   &\multicolumn{3}{|c|}{Time} &\multicolumn{3}{|c|}{States} &\multicolumn{3}{|c|}{ Paths }\\
								&	&\gsetxt{} & L\#		&SL		&\gsetxt{}	& L\# & SL&\gsetxt{}	& L\# 	& SL\\
   \hline
    \multirow{3}{*}{LinkedList }			&3	& 0.91	& 1.21	& 0.69	& 2465	& 2844	& 99		&1656	& 1269	& 25\\
   		 						& 4	& 2.92	& 3.35	& 0.91	& 25774	& 29977	& 155	&17485	& 13550	& 39\\
   								& 5	& 20.78	& 19.47	& 1.59	& 341164	& 400296	& 223	&232743	& 181849	& 56\\
								& 6	& 280.56	& 299.19	& 2.36	&5447980	&6437201	& 303	&3731094	&2933027	& 76\\
    \hline
    \multirow{3}{*}{BinarySearchTree }	& 1	& 0.26	& 0.28	& 0.36	& 19		& 23		& 29		& 6		& 6		& 6\\
   		 						& 2	& 0.83	& 1.28	& 0.93	& 143	& 143	& 145	& 43		& 42		& 33\\
   								& 3	& 20.63	& 25.55	& 4.03	& 1953	& 1703	& 1485	& 515	& 515	& 328\\
    \hline
      \multirow{3}{*}{TreeMap}			& 1	& 0.47	& 0.52	& 0.77	& 65		& 70		& 215	& 11		& 11		& 11\\
   		 						& 2	& 8.99	& 9.73	& 4.72	& 1009	& 942	& 3219	& 127	& 122	& 73\\
   								& 3	& -		& -		& 145.56	& -		& -		& 78695	& -		& -		& 887\\

@@

The results of the experiments are presented in Table
\ref{tab:results}. Each row contains the results for the specified $k$-bound 
for each artifact evaluated. The columns show the total
wall clock time in seconds, states explored, and paths generated for
each of the algorithms. The headings \gsetxt{}, L\#, and SL correspond
to the Generalized Symblic Execution, lazier\#, and \symtxt{}
algorithms, respectively. Note that we do no present the results for
lazier algorithm since they are almost identical to that of lazier\#
for our examples. Table entries containing - correspond to test runs
that exceeded the allotted time bounds. Each trial is time bounded to
10 mins.

In most tests, run times strongly favor \symtxt{}, especially for
large $k$-values. Performance improvement ranges from $4.8x$ for
BinarySearchTree at $k$=3, to $118x$ for LinkedList at $k$=6. In fact,
for certain $k$-bounds a number of experiments completed exploration
using \symtxt{} in a few seconds whereas \gsetxt{} or lazier\# were
unable to do, e.g., BinarySearchTree for $k$=4 and LinkedList for
$k$=8. These results demonstrate the potential of using \symtxt{} as
an approach to handle data input of complex types. The lower runtimes
show that we can prove properties using \symtxt{} faster compared to
\gsetxt{}.

@@

Path counts likewise substantially favor \symtxt{}. The number of
paths explored by \symtxt{} is strictly less than or equal the number
of paths explored by \gsetxt{} for all test cases. We can use this
result to do more efficient test case generation. From a path
generated we can use the solutions provided by the constraint solver
to instantiate a set of concrete heaps to provide as test input. These
can provide a smaller test suite to achieve control-flow path coverage
compared to GSE and Lazier\#.

@@

Curiously, the pattern established with run-times and path counts does
not hold true for state counts. State counts can vary between
algorithms, because states in JPF represent points of nondeterminism,
and each algorithm differs in this regard. For example, \gsetxt{} has
additional points of nondeterminism during field reads, but in
contrast references compares are completely deterministic. Thus, in
example programs with large numbers reference compares, such as
TreeMap, state counts for lazier\# and \symtxt{} may exceed those for
\gsetxt{}. The larger state count, however, does not contribute to the
runtime of the technique.

@@

In summary, it appears that the benefits of avoiding
initialization-based nondeterminism more than outweigh the increased
complexity in the constraints due to the advances made in SMT
solvers. In fact, \symtxt{} can analyze certain types of programs with
orders of magnitude greater efficiency than with \gsetxt{} or
lazier\#, while covering exactly the same control flow paths.

>>>>

%\begin{figure*}[t]
\begin{table*} [t]
  \centering
  \scalebox{0.88}{\begin{tabular}{| c | c | r | r | r | r | r | r | r | r | r |}
  \hline
   \multirow{2}{*}{Method }&\multirow{2}{*}{ $k$ }
   &\multicolumn{3}{|c|}{Time (seconds)} &\multicolumn{3}{|c|}{States} &\multicolumn{3}{|c|}{ Paths }\\
								&	&\gsetxt{} & L\#		&SL		&\gsetxt{}	& L\# & SL&\gsetxt{}	& L\# 	& SL\\
   \hline
    \multirow{3}{*}{LinkedList }			&3	& 0.91	& 1.21	& 0.69	& 2465	& 2844	& 99		&1656	& 1269	& 25\\
   		 						& 4	& 2.92	& 3.35	& 0.91	& 25774	& 29977	& 155	&17485	& 13550	& 39\\
   								& 5	& 20.78	& 19.47	& 1.59	& 341164	& 400296	& 223	&232743	& 181849	& 56\\
								& 6	& 280.56	& 299.19	& 2.36	&5447980	&6437201	& 303	&3731094	&2933027	& 76\\
    \hline
    \multirow{3}{*}{BinarySearchTree }	& 1	& 0.26	& 0.28	& 0.36	& 19		& 23		& 29		& 6		& 6		& 6\\
   		 						& 2	& 0.83	& 1.28	& 0.93	& 143	& 143	& 145	& 43		& 42		& 33\\
   								& 3	& 20.63	& 25.55	& 4.03	& 1953	& 1703	& 1485	& 515	& 515	& 328\\
    \hline
      \multirow{3}{*}{TreeMap}			& 1	& 0.47	& 0.52	& 0.77	& 65		& 70		& 215	& 11		& 11		& 11\\
   		 						& 2	& 8.99	& 9.73	& 4.72	& 1009	& 942	& 3219	& 127	& 122	& 73\\
   								& 3	& -		& -		& 145.56	& -		& -		& 78695	& -		& -		& 887\\

@@

The results of the experiments are presented in Table
\ref{tab:results}. Each row contains the results for the specified $k$-bound 
for each artifact evaluated. The columns show the total
wall clock time in seconds, states explored, and paths generated for
each of the algorithms. The headings \gsetxt{}, L\#, and SL correspond
to the Generalized Symblic Execution, lazier\#, and \symtxt{}
algorithms, respectively. Note that we do no present the results for
the lazier algorithm since they are almost identical to that of lazier\#
for our examples. Table entries containing '-' correspond to test runs
that exceeded the allotted time bound of 10 minutes.

In most tests, run times strongly favor the \symtxt{} algorithm, especially for
large $k$-values. Performance improvement ranges from $4.8x$ for
BinarySearchTree at $k$=3, to $118x$ for LinkedList at $k$=6. In fact,
for certain $k$-bounds a number of experiments completed exploration
using the \symtxt{} algorithm in a few seconds whereas \gsetxt{} or lazier\# were
unable to finish, e.g., BinarySearchTree for $k$=4 and LinkedList for
$k$=8. These results demonstrate the potential of using the \symtxt{} algorithm 
to handle data input for complex types. The lower running times
show that we can prove properties using the \symtxt{} algorithm faster than
\gsetxt{}.

@@

Path counts likewise substantially favor the \symtxt{} algorithm. The number of
paths explored by the algorithm is strictly less than or equal to the number
of paths explored by \gsetxt{} for all test cases. We can use this
result to do more efficient test case generation. From a path
generated we can use the solutions provided by the constraint solver
to instantiate a set of concrete heaps to provide as test input. These
can provide a smaller test suite to achieve control-flow path coverage when
compared to GSE and Lazier\#.

@@

Curiously, the pattern established with run-times and path counts does
not hold true for state counts. State counts can vary between
algorithms because states in JPF represent points of nondeterminism,
and each algorithm differs in this regard. For example, \gsetxt{} has
additional points of nondeterminism during field reads, but in
contrast reference compares are completely deterministic. Thus, in
example programs with large numbers of reference compares, such as
TreeMap, state counts for lazier\# and the \symtxt{} algorithm may exceed those for
\gsetxt{}. The larger state count, however, does not contribute to the
runtime of the technique.

@@

In summary, the benefits of avoiding
initialization-based non-determinism outweigh the increased
complexity in the constraints due to the advances made in SMT
solvers in these examples. In fact, the \symtxt{} algorithm can analyze certain types of programs with
orders of magnitude greater efficiency than with \gsetxt{} or
lazier\#, while covering exactly the same control flow paths.

### 2014-11-13 tex/results.tex
Finished Eval

\section{Evaluation}
\label{sec:eval}
In this section we evaluate the effectiveness of the \symtxt{}
approach at analyzing programs with unbounded complex input data in
the context of symbolic execution. To perform our evaluation
we implement a prototype of the \symtxt{} technique and evaluate its efficiacy on
three Java data structure artifacts. We compare the results of the
\symtxt{} approach with other state-of-the-art symbolic execution techniques 
that support analysis of program with complex input data types.

@@

The prototype of  our~\symtxt{} approach is implemented as a new extension to SPF.
In addition to the operations presented in the Javalite
semantics (see \secref{sec:precise}), the implementation contains 
support for operations over
integers. Examples of bytecode implemented in SPF to support \symtxt{} 
algorithm include:
(i) field access bytecodes, e.g., GETFIELD and ALOAD, to implement the 
field-access and summarize rules
in~\figref{fig:symInit} and~\figref{fig:fHeap}(a) respectively, (ii) the field
write bytecodes, e.g., PUTFIELD and ASTORE, to implement the
field-write rules in~\figref{fig:fHeap}(a), and (iii) equality checking of
references bytecodes, IF\_ACMPEQ and IF\_ACMPNE, to implement the
equals references-true and references-false rules in~\figref{fig:eqs}.
We also implement linear arithmetic operations on integers, and integer
comparison operations. We are currently working on adding support for
floating point operations, arrays, and bit-operations.  We use the
\texttt{jConstraints} library which provides support for handling heap
constraints~\cite{ase2014-ghilrr,jpf2014-dghirr}. It uses the z3
solver~\cite{deMouraBjorner08Z3} as the underlying constraint solver. We
leverage the incremental solving feature in z3 in our experiments. We
also implement other optimizations that cache and re-use solutions to
previous constraints. These optimizations are an important part of our
implementation because only small portions of the heap constraint
changes from one state to another.

@@

The evaluation is performed using three Java data structure artifacts 
(a) Linked List,
(b) Binary Search Tree, and (c) Tree Map. 
\sjp{next sentence is bumpy}
These data structure
programs are often used to evaluate symbolic execution algorithms
that analyze software using arbitrary heap inputs. The \emph{independent} variable in our
study is the algorithm for initializing symbolic references and the
$k$-bound. The initialization algorithms that we vary in our study
are: \symtxt{}, \gsetxt{}, lazier, and lazier\#. We also vary the
$k$-bound for each of the algorithms. Note that $k$-bounding bounds the
length of a reference chain from the root of the heap~\cite{Deng:2006}.
We select three dependent variables and measures for our study: (i)
time, (ii) states explored, and (iii) paths generated. The \emph{time}
is the total time taken by each algorithm to explore the symbolic
execution tree as well as total constraint solving time. The
\emph{states explored} represents the number of nodes in the symbolic
execution tree, and the \emph{paths generated} represents 
the number of unique paths in the symbolic execution tree.

>>>>

\section{Evaluation}
\label{sec:eval}
In this section we evaluate the effectiveness of the \symtxt{}
algorithm at analyzing programs with unbounded complex input data in
the context of symbolic execution. To perform our evaluation
we implement a prototype of the algorithm and evaluate its efficiacy on
three Java data structure artifacts. We compare the results of the
\symtxt{} approach with other state-of-the-art symbolic execution techniques 
that support analysis of programs with complex input data types.

@@

The prototype of  our~\symtxt{} algorithm is implemented as a new extension to SPF.
In addition to the operations presented in the Javalite
semantics (see \secref{sec:precise}), the implementation contains 
support for operations over
integers including symbolic representations. Examples of bytecode implemented in SPF to support the \symtxt{} algorithm
include:
(i) field access bytecodes, e.g., GETFIELD and ALOAD, to implement the 
field-access and summarize rules
in~\figref{fig:symInit} and~\figref{fig:fHeap}(a) respectively, (ii) the field
write bytecodes, e.g., PUTFIELD and ASTORE, to implement the
field-write rules in~\figref{fig:fHeap}(a), and (iii) equality checking of
references bytecodes, IF\_ACMPEQ and IF\_ACMPNE, to implement the
equals references-true and references-false rules in~\figref{fig:eqs}.
We also implement linear arithmetic operations on integers, and integer
comparison operations. We are currently working on adding support for
floating point operations, arrays, and bit-operations.  We use the
\texttt{jConstraints} library which provides support for handling heap
constraints~\cite{ase2014-ghilrr,jpf2014-dghirr}. It uses the z3
solver~\cite{deMouraBjorner08Z3} as the underlying constraint solver. We
leverage the incremental solving feature in z3 in our experiments. We
also implement other optimizations that cache and re-use solutions to
previous constraints. These optimizations are an important part of our
implementation because only small portions of the heap constraint
changes from one state to another.

@@

The evaluation is performed using three Java data structure artifacts 
(i) Linked List,
(ii) Binary Search Tree, and (iii) Tree Map. 
These are the typical data structures for evaluating symbolic execution algorithms with heap initialization. The \emph{independent} variable in our
study is the algorithm for initializing symbolic references and the
$k$-bound. The initialization algorithms that we vary in our study
are: \symtxt{}, \gsetxt{}, lazier, and lazier\#. We also vary the
$k$-bound for each of the algorithms. Note that $k$-bounding bounds the
length of a reference chain from the root of the heap~\cite{Deng:2006}.
We select three dependent variables and measures for our study: (i)
time, (ii) states explored, and (iii) paths generated. The \emph{time}
is the total time taken by each algorithm to explore the symbolic
execution tree as well as total constraint solving time. The
\emph{states explored} represents the number of nodes in the symbolic
execution tree, and the \emph{paths generated} represents 
the number of unique paths in the symbolic execution tree.
